Cosmology with photometric weak lensing surveys: constraints with redshift tomography of convergence peaks and moments

### 2016-09-22 photoz_paper/paper.tex
implemented referee comments

Several different examples of these non-Gaussian statistics, and their cosmological information content, have been studied in the past as well (see \citep{MinkJan,PeaksJan,NG-Marian,NG-Jain1,NG-Jain2,NG-Jain3,NG-Refregier,NG-Dietrich} for a non-comprehensive list). The constraining power of weak lensing power spectra with the addition of redshift tomography information have been extensively investigated in the literature (see e.g. \citep{SongKnox,FangHaiman07,Huterer2006}). In this work we concentrate on the constraining power of a subset of non-Gaussian statistics, combined with redshift tomography in an LSST-like survey. Previous work on redshift tomography with weak lensing Minkowski functionals is also present in the literature \citep{MinkJan}.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating simulated shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with parameters $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$ \citep{WMAP9,PlanckCosmo}. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses out to redshift $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the ordinary differential lens equations up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from large scale structure, assuming the non-lensed shape is a circle.

@@

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropic, the expectation value $\langle\rangle$, for each realization $r$, is taken over all modes $\pmb{\ell}$ with the same magnitude $\ell=\vert\pmb{\ell}\vert$. The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the standard deviation over all pixels. The set of nine moments $\pmb{\mu}(\bar{z}_b)$ is defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}): 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c) .
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity. In the next section, we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from simulated observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$ and $\pmb{\mu}(\bar{z}_b)$.

@@

Figure \ref{fig:photozbias} shows the effect of ignoring photo-$z$ errors on parameter constraints. To evaluate this effect we construct different simulated observations, with and without photo-$z$ errors, and compare the results of the parameter fit according to equation (\ref{meth:parestimate}). Using our simulation suite, we construct 20 simulated observations: the summary statistic in each observation is calculated by taking the mean of a random sample of $N_{\rm FOV}=1600$ realizations of the summary statistic in the fiducial cosmology (randomly chosen among the ensemble of $N_r=16000$ that are available in the ensemble). The estimated covariance matrix $\bbh{C}$ is scaled by a factor $N_{\rm FOV}$ to take into account the construction process of the simulated observations. This procedure allows us to forecast the results an LSST-like survey would obtain.

@@

Figure \ref{fig:photozbias} quantifies the effect of uncorrected photo-$z$ errors on the $(\Omega_m,w)$ constraints. Because the stochastic nature of the observations, parameter values $\bbh{p}$ estimated from equation (\ref{meth:parestimate}) are affected by statistical fluctuations. In Figure \ref{fig:photozbias} we show 20 random draws from the probability distribution of $\delta \bbh{p}=\bbh{p}_{\rm photo-z}-\bbh{p}_{\rm no-photo-z}$. We can conclude that the $\bbh{p}$ estimator is biased if $\langle\delta\bbh{p}\rangle\neq 0$. Figure \ref{fig:photozbias} clearly show that photo-$z$ errors cause parameter biases at more than $1\sigma$ significance level when using the power spectrum and the moments, while no bias is observed for the peak count statistic within its 68\% confidence region. We also observe that photo-$z$ errors bias the constraints in slightly different directions, leaving open the possibility of identifying and correcting this bias through self-calibration techniques.

>>>>

Several different examples of these non-Gaussian statistics, and their cosmological information content, have been studied in the past as well (see \citep{MinkJan,PeaksJan,NG-Marian,NG-Jain1,NG-Jain2,NG-Jain3,NG-Refregier,NG-Dietrich} for a non-comprehensive list). The constraining power of weak lensing power spectra with the addition of redshift tomography information have been extensively investigated in the literature (see e.g. \citep{SongKnox,FangHaiman07,Huterer2006}). In this work we concentrate on the constraining power of a subset of non-Gaussian statistics, combined with redshift tomography in an LSST-like survey. {\bf \color{red} \citep{MartinetPeaksTomo} investigated the cosmological constraining power of shear peaks tomography}. Previous work on redshift tomography with weak lensing Minkowski functionals is also present in the literature \citep{MinkJan}.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating simulated shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with parameters $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$ \citep{WMAP9,PlanckCosmo}. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle.

{\bf \color{red} The largest mode observed in our $N$--body simulations corresponds to a wavenumber of $k_b\approx1/L_b\approx 0.004 h{\rm Mpc}^{-1}$. For the sake of recovering cosmological information from WL, this limitation does not create a concern, as several authors (see \citep{FangHaiman07} for example) have shown that modes above $L_b$ contribute very little to parameter constraints. 
Moreover, the purpose of this work is to estimate the parameter constraints achievable in a weak lensing analysis incorporating tomography, not to produce simulations accurate enough for analyzing the data set that will be available from LSST and other surveys a decade hence. To analyze the datasets that these surveys will produce, mode couplings between large and small scales, which can cause effects such as super sample covariance \citep{Sato12,SSC1,SSC2}, will need to be included. Baryonic effects will need to be included as well. Larger and more accurate $N$--body simulation techniques are currently under development in the community for this purpose \citep{Qcontinuum,HACC}.}

The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses out to redshift $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the ordinary differential lens equations up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from large scale structure, assuming the non-lensed shape is a circle.

@@

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropic, the expectation value $\langle\rangle$, for each realization $r$, is taken over all modes $\pmb{\ell}$ with the same magnitude $\ell=\vert\pmb{\ell}\vert$. {\bf \color{red} Given the fact that our simulation box is small, and we are ignoring non--linear couplings between large and small scale modes, we are likely underestimating the $\kappa$ power spectrum when performing ensemble averages based on a single $N$--body box. \citep{NbodySample} estimated the effect of a varying box size on the 3D matter power spectrum, for boxes up to 512Mpc$/h$ in size and found the variations to be small compared to their sample variance, on spatial wavenumbers up to $k\sim 0.3h{\rm Mpc}^{-1}$. }

The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the standard deviation over all pixels. The set of nine moments $\pmb{\mu}(\bar{z}_b)$ is defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}): 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c) .
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity. In the next section, we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from simulated observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$ and $\pmb{\mu}(\bar{z}_b)$. {\bf \color{red} Concerns might arise on the accuracy with which our simulations measure the summary statistics mentioned above, given the small box size and the fact that we recycle a single $N$--body box for building our simulated sample. \citep{PetriVariance} studied the dependence of the power spectrum and peak counts sample means as a function of the number of independent $N$--body boxes and found that the variations are less than 10\% in most cases, except for the small scale power spectrum and the highest $\kappa$ peaks, for which the variations are less than 20\%. }

@@

Figure \ref{fig:photozbias} shows the effect of ignoring photo-$z$ errors on parameter constraints. To evaluate this effect we construct different simulated observations, with and without photo-$z$ errors, and compare the results of the parameter fit according to equation (\ref{meth:parestimate}). Using our simulation suite, we construct 20 simulated observations: the summary statistic in each observation is calculated by taking the mean of a random sample of $N_{\rm FOV}=1600$ realizations of the summary statistic in the fiducial cosmology (randomly chosen among the ensemble of $N_r=16000$ that are available in the ensemble). The estimated covariance matrix $\bbh{C}$ is scaled by a factor $N_{\rm FOV}$ to take into account the construction process of the simulated observations. This procedure allows us to forecast the results an LSST-like survey would obtain. { \bf \color{red} We stress that, because of the small size of our simulation box the covariance estimate that we obtain is likely not accurate enough to produce constraints from LSST data. Full treatment of observables covariance matrices, along with larger $N$--body simulations and SSC effects will be investigated in future work.}

@@

Figure \ref{fig:photozbias} quantifies the effect of uncorrected photo-$z$ errors on the $(\Omega_m,w)$ constraints. Because the stochastic nature of the observations, parameter values $\bbh{p}$ estimated from equation (\ref{meth:parestimate}) are affected by statistical fluctuations. In Figure \ref{fig:photozbias} we show 20 random draws from the probability distribution of $\delta \bbh{p}=\bbh{p}_{\rm photo-z}-\bbh{p}_{\rm no-photo-z}$. We can conclude that the $\bbh{p}$ estimator is biased if $\langle\delta\bbh{p}\rangle\neq 0$. Figure \ref{fig:photozbias} clearly show that photo-$z$ errors cause parameter biases at more than $1\sigma$ significance level when using the power spectrum and the moments, while no bias is observed for the peak count statistic within its 68\% confidence region. {\bf \color{red} Peak histogram shapes are more robust to this kind of systematic effect since the peak locations are determined by the information coming from neighboring galaxies, while the photo-z errors have no spatial correlation. Photo-z errors are more likely to alter point estimates of the $\kappa$ distribution and larger scale correlations which affect the power spectrum.}

We also observe that photo-$z$ errors bias the constraints in slightly different directions, leaving open the possibility of identifying and correcting this bias through self-calibration techniques.
